---
output: 
      html_document:
            css: hospitals-style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width="100%", out.height="100%")

pacman::p_load(dplyr, glue, reactable, highcharter, htmltools, dataui)

react_tab_final <- readr::read_rds(glue("{rprojroot::find_rstudio_root_file()}/Indiana-COVID-19-Tracker/data/hosp-react-tab.rds"))
mort_staff_admiss_skew_dat <- readr::read_rds(glue("{rprojroot::find_rstudio_root_file()}/Indiana-COVID-19-Tracker/data/hosp-msas-line.rds"))
goog_key <- Sys.getenv("GOOG_API")


```



```{r state-msa-style-funs, include=FALSE}
# Shared tooltip formatting function
# this.points is your data (but not your date column for some reason), zero indexed
# series.name is the series 'name' you entered below
# y is value for the column you entered for 'y' for that series
# this.x is the variable you entered for the xAxis
fntltp <- JS("function(){
              return '<b>' + this.x + '</b>' + '  (Rolling Averages)<br>' +
              '<i>' + this.points[0].series.name + '</i>' + ' : ' + this.points[0].y.toLocaleString(undefined, {maximumFractionDigits: 1, style: 'percent'}) + '<br>' +
              '<i>' + this.points[1].series.name + '</i>' + ' : ' + this.points[1].y.toLocaleString(undefined, {maximumFractionDigits: 1, style: 'percent'}) + '<br>' +
              '<i>' + this.points[2].series.name + '</i>' + ' : ' + this.points[2].y.toFixed(1) + '<br>' +
              '<i>' + this.points[3].series.name + '</i>' + ' : ' + this.points[3].y.toFixed(2);
}")

# fun for yaxis labels
yaxis_fmt <- JS("function(){
      return this.value.toLocaleString(undefined, {maximumFractionDigits: 0, style: 'percent'});
   }")

# funs for data labels
fmt_labs_perc <- JS("function(){
               return this.point.y.toLocaleString(undefined, {maximumFractionDigits: 1, style: 'percent'})
}")
fmt_labs_adm <- JS("function(){
               return this.point.y.toFixed(1)
}")
fmt_labs_skew <- JS("function(){
               return this.point.y.toFixed(2)
}")


theme_hosp_msa <- hc_theme(
      chart = list(
            backgroundColor = "hsl(0, 0%, 5%)"
      ),
      yAxis = list(
            title = list(text = NULL),
            gridLineWidth = 0
      ),
      xAxis = list(
            labels = list(
                  style = list(fontSize = "14px",
                               color = "white")
            )
      ),
      legend = list(
            layout = "horizontal",
            align = "center",
            verticalAlign = "top",
            itemStyle = list(color = "white")
      ),
      caption = list(
            align = "right",
            style = list(color = "white")
      )
)

```



## **State Hospital Mortality, Staffing Shortages, Admissions**

```{r state-msa-line, message=FALSE, echo=FALSE}
state_msa_line <- highchart() %>%
   # Couldn't get HC to hand date class so using ordered factor for date var
   hc_xAxis(categories = levels(mort_staff_admiss_skew_dat$date),
            # label every <step> days
            labels = list(step = 2)) %>%
   
   #Hospital Mortality Rate
   # nid is for relative arg, syncs series to each other
   hc_add_yAxis(nid = 1L,
                labels = list(style = list(color = "white"),
                              formatter = yaxis_fmt),
                # HC calcs funky axis breaks sometimes; tickamounts = num_breaks
                min = 0.10, max = 0.20, tickAmount = 6,
                # first label removed because it's distorted in this facetted chart format
                # relative syncs the tooltip to chart 3 (could also choose 2 probably)
                showLastLabel = FALSE, relative = 1) %>%
   hc_add_series(name = 'Hospital Mortality Rate', data = mort_staff_admiss_skew_dat,
                 # increased lineWidth to highlight mort rate
                 type = "line", lineWidth = 2.5,
                 hcaes(x = date, y = mean_mort_rate),
                 color = prismatic::clr_lighten("#a88814", shift = 0.3),
                 # yAxis is the axis id (O indexed)
                 yAxis = 0, tooltip = list(stickOnContact = FALSE),
                 # states has various js-like effects type features
                 # turned off effect of series being highlighted on hover
                 states = list(inactive = list(enabled = FALSE)),
                 # add label to latest data value
                 dataLabels = list(
                    enabled = TRUE,
                    backgroundColor = "white",
                    color = "black",
                    formatter = fmt_labs_perc,
                    filter = list(
                       operator = "==",
                       property = "index",
                       # 0-indexed
                       value = nrow(mort_staff_admiss_skew_dat)-1
                    ),
                    # moves label above point
                    y = -15)) %>% 
   
   # Hospitals with Staffing Shortages
   hc_add_yAxis(nid = 2L,
                labels = list(style = list(color = "white"),
                              formatter = yaxis_fmt),
                min = min(mort_staff_admiss_skew_dat$roll_mean_staff_short_perc) - 0.02,
                max = max(mort_staff_admiss_skew_dat$roll_mean_staff_short_perc) + 0.02,
                tickAmount = 4, showLastLabel = FALSE, relative = 1) %>%
   hc_add_series(name = 'Hospitals with Staffing Shortages', data = mort_staff_admiss_skew_dat,
                 type = "line", hcaes(x = date, y = roll_mean_staff_short_perc), color = "#a8149e",
                 yAxis = 1, tooltip = list(stickOnContact = FALSE),
                 
                 states = list(inactive = list(enabled = FALSE)),
                 lineWidth = 1, opacity = 1, marker = list(radius = 3),
                 dataLabels = list(
                    enabled = TRUE,
                    backgroundColor = "white",
                    color = "black",
                    formatter = fmt_labs_perc,
                    filter = list(
                       operator = "==",
                       property = "index",
                       value = nrow(mort_staff_admiss_skew_dat)-1
                    ),
                    y = -15)) %>%
   
   # Hospital Admissions
   hc_add_yAxis(nid = 3L,
                labels = list(style = list(color = "white")),
                tickAmount = 6, gridLineWidth = 0, 
                showLastLabel = FALSE, showFirstLabel = TRUE, relative = 4) %>% 
   hc_add_series(name = 'Hospital Admissions', data = mort_staff_admiss_skew_dat,
                 type = "line", hcaes(x = date, y = roll_mean_admiss), color = '#1e14a8',
                 yAxis = 2, tooltip = list(stickOnContact = FALSE),
                 states = list(inactive = list(enabled = FALSE)),
                 lineWidth = 1, opacity = 1, marker = list(radius = 2),
                 dataLabels = list(
                    enabled = TRUE,
                    backgroundColor = "white",
                    color = "black",
                    formatter = fmt_labs_adm,
                    filter = list(
                       operator = "==",
                       property = "index",
                       value = nrow(mort_staff_admiss_skew_dat)-1
                    ),
                    y = -15)) %>% 
   
   # Age Skew in Admissions
   hc_add_yAxis(nid = 4L,
                labels = list(style = list(color = "white")),
                min = min(mort_staff_admiss_skew_dat$skewness) - 0.05,
                max = max(mort_staff_admiss_skew_dat$skewness) + 0.05,
                tickAmount = 4, gridLineWidth = 0, 
                showLastLabel = FALSE, relative = 1) %>% 
   hc_add_series(name = 'Age Skewness in Admissions', data = mort_staff_admiss_skew_dat,
                 type = "line", hcaes(x = date, y = skewness), color = '#6814a8',
                 yAxis = 3, tooltip = list(stickOnContact = FALSE),
                 states = list(inactive = list(enabled = FALSE)),
                 lineWidth = 1, opacity = 1, marker = list(radius = 2),
                 dataLabels = list(
                    enabled = TRUE,
                    backgroundColor = "white",
                    color = "black",
                    formatter = fmt_labs_skew,
                    filter = list(
                       operator = "==",
                       property = "index",
                       value = nrow(mort_staff_admiss_skew_dat)-1
                    ),
                    y = -15)) %>% 
   
   # shared creates 1 tooltip box for all series and syncs them
   hc_tooltip(formatter = fntltp,
              shared = TRUE) %>% 
   hc_caption(text = "Sources: Regenstrief Institute, Department of Health and Human Services") %>% 
   hc_add_theme(theme_hosp_msa) %>% 
   hc_size(height = 900)

state_msa_line

```

This chart shows Indiana's hospital mortality rate and unique hospital admissions data collected by the [Regenstrief Institute](https://www.regenstrief.org/covid-dashboard/). The hospital staffing [data](https://healthdata.gov/dataset/covid-19-reported-patient-impact-and-hospital-capacity-state-timeseries) comes from the Department of Health and Human Services (HHS). If hospitals across the state become overwhelmed, resources will get rationed and patient outcomes may worsen. Therefore, a relationship between hospital admissions, staffing shortages, and the age make-up of admissions with hospital mortality rate seems like a reasonable assumption. If any of these three factors do influence hospital mortality, the effects aren't likely to manifest for days or maybe even a couple weeks. Further details on the data and calculations can be found in the [changelog](https://github.com/ercbk/Indiana-COVID-19-Tracker/blob/master/NEWS.md).  

   * *Hospital Mortality Rate* is the ratio of deaths of hospitalized patients and unique hospital admissions. Each rate is calculated over a rolling, 14-day window.  
   * *Hospitals with Staffing Shortages* is the percent of reporting hospitals that have reported staffing shortages for that day. That percentage is averaged over a rolling, seven day window.  
   * *Hospital Admissions* is the daily total of unique individuals admitted to a hospital on the that day. Those daily counts are averaged over a rolling, seven day window.  
   * *Age Skewness in Admissions* is a measurement of the age make-up of the hospital admissions data. It's calculated over a rolling, 14 day window (same as the hospital mortality rate). The larger the value, the older the group of admissions is on average. Between 0.5 and 1, the group of admissions is moderately skewed towards older individuals. Greater than 1, the group of admissions is highly skewed towards older individuals.   
   
The latest data values for each variable are labeled. Panning across the chart will display values for all variables on that day. Each line series can be selected or de-selected by clicking on names in the legend.

\
\
\



```{r react_funs, include=FALSE}

# reformat to Month day
react_data_date <- as.character(format(max(react_tab_final$end_date), "%B %d"))
# # header function to create tooltip
with_tooltip <- function(value, tooltip) {
      span(title = tooltip, value)
}


# Calls Google Maps API when row is expanded
row_details <- function(index) {
      dat <- react_tab_final %>% 
            slice(index)
      
      goog_api_params <- dat %>% 
            distinct(hospital_name, city_zip) %>%
            # extracts 1 or 2 words before the comma
            mutate(city = stringr::str_extract(city_zip, pattern = "[:alpha:]* [:alpha:]*(?=,)|[:alpha:]*(?=,)")) %>% 
            mutate_at(vars(hospital_name, city), stringr::str_replace_all, pattern = " ", replacement = "+") %>% 
            select(-city_zip)
      
      goog_map_url <- glue::glue("https://www.google.com/maps/embed/v1/place?key={goog_key}&q={goog_api_params$hospital_name[[1]]},{goog_api_params$city[[1]]},IN")
      
      details <- tagList(
            div(
                  tags$br(),
                  tags$iframe(
                        src=goog_map_url,
                        width="400",
                        height="300",
                        frameborder="0",
                        style="border:0; margin-left:10px",
                        allowfullscreen="",
                        tabindex="0",
                        marginwidth="50"
                  )
            )
      )
      return(details)
}

# Sparkline column functions
icuBeds_spark <- function(...) {
      colDef(
            name = "Average Daily ICU Beds Occupied by COVID Patients",
            minWidth = 200,
            cell = dui_for_reactable(
                  dui_sparkline(
                        # name of the list of all the inner pairwise lists (i.e. outmost list)
                        data = htmlwidgets::JS("cellInfo.value.avg_covid_icu_10k_list"),
                        # y-axis value from each inner pairwise list
                        valueAccessor = htmlwidgets::JS("(d) => d.avgCovIcuTenK[0]"),
                        renderTooltip = htmlwidgets::JS(
                              htmltools::HTML(
                                    "function (_ref) {
                   var datum = _ref.datum;
                   // div around two spans makes horiz. date-value tooltip
                   // react element format (type, style, function)
                   return React.createElement(
                     'div',
                     null,
                     datum.endDate && React.createElement(
                        'span',
                        {style: {
                             backgroundColor: 'black', color: 'white',
                             padding: '3px', margin: '0px 4px 0px 0px', textAlign: 'center'
                           }},
                        // splits date at '-', removes first elt, pastes remainder elts with '/'
                        datum.endDate[0].split('-').slice(1).join('/')
                     ),
                     React.createElement(
                        'span',
                        {style: {
                        fontWeight: 'bold', fontSize: '1.1em',
                        padding: '1px'
                        }},
                        // undefined is the region (ie locale), {options}, maximumFractionDigits rounds up
                        datum.y ? datum.y.toLocaleString(undefined, {maximumFractionDigits: 1}) : '--'
                     )
                   );
                  }"
                              )
                        ),
                        components = list(
                              dui_sparklineargradient(
                                    id = "bar_gradient1",
                                    to = icu_light,
                                    from = icu_dark
                              ),
                              dui_sparkbarseries(fill = "url(#bar_gradient1)",
                                                 stroke = "#e57647"),
                              dui_sparklineseries(stroke = "#e57647",
                                                  strokeWidth = 0.8)
                        )
                  )
            )
      )
} 

hospBeds_spark <- function(...) {
      colDef(
            name = "Average Daily Hospital Beds Occupied by COVID Patients",
            minWidth = 200,
            cell = dui_for_reactable(
                  dui_sparkline(
                        # name of the list of all the inner pairwise lists (i.e. outmost list)
                        data = htmlwidgets::JS("cellInfo.value.avg_covid_hosp_10k_list"),
                        # y-axis value from each inner pairwise list
                        valueAccessor = htmlwidgets::JS("(d) => d.avgCovHospTenK[0]"),
                        renderTooltip = htmlwidgets::JS(
                              htmltools::HTML(
                                    "function (_ref) {
                   var datum = _ref.datum;
                   // div around two spans makes horiz. date-value tooltip
                   // react element format (type, style, function)
                   return React.createElement(
                     'div',
                     null,
                     datum.endDate && React.createElement(
                        'span',
                        {style: {
                             backgroundColor: 'black', color: 'white',
                             padding: '3px', margin: '0px 4px 0px 0px', textAlign: 'center'
                           }},
                        // splits date at '-', removes first elt, pastes remainder elts with '/'
                        datum.endDate[0].split('-').slice(1).join('/')
                     ),
                     React.createElement(
                        'span',
                        {style: {
                        fontWeight: 'bold', fontSize: '1.1em',
                        padding: '1px'
                        }},
                        // undefined is the region (ie locale), {options}, maximumFractionDigits rounds up
                        datum.y ? datum.y.toLocaleString(undefined, {maximumFractionDigits: 1}) : '--'
                     )
                   );
                  }"
                              )
                        ),
                        components = list(
                              dui_sparklineargradient(
                                    id = "bar_gradient2",
                                    to = hosp_light,
                                    from = hosp_dark
                              ),
                              dui_sparkbarseries(fill = "url(#bar_gradient2)",
                                                 stroke = "#47c8e5"),
                              dui_sparklineseries(stroke = "#47c8e5",
                                                  strokeWidth = 0.8)
                        )
                  )
            )
      )
}

```


```{r react-styling, include=FALSE}

# palette for heatmap columns
perc_pal <- scales::col_numeric(
      palette =  colorspace::sequential_hcl(100, "ag_Sunset"),
      domain = c(0, 1.00),
      na.color = "hsl(0, 0%, 10%)",
      reverse = TRUE
)

# palettes for sparklines
hosp_light <- prismatic::clr_lighten("#47c8e5", shift = 0.60)
hosp_dark <- prismatic::clr_darken("#47c8e5", shift = 0.60)
icu_light <- prismatic::clr_lighten("#d5e547", shift = 0.60)
icu_dark <- prismatic::clr_darken("#d5e547", shift = 0.60)

react_theme = reactableTheme(
      color = "hsl(233, 9%, 87%)",
      backgroundColor = "hsl(0, 0%, 5%)",
      borderColor = "hsl(233, 9%, 22%)",
      stripedColor = "hsl(233, 12%, 22%)",
      highlightColor = "hsl(233, 12%, 24%)",
      inputStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
      selectStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
      pageButtonHoverStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
      pageButtonActiveStyle = list(backgroundColor = "hsl(233, 9%, 28%)"))

```

## **Local Hospital Capacity**

```{r reactable, message=FALSE, echo=FALSE}


hosp_react_tab <- reactable(
      data = react_tab_final %>% 
            select(hospital_name, county_name,
                   sev_day_icu_perc_occup, sev_day_hosp_perc_occup,
                   avgCovIcuTenKList, avgCovHospTenKList),
      borderless = TRUE,
      searchable = TRUE,
      style = list(fontSize = "14px"),
      # expand when details button clicked
      onClick = "expand",
      # rows per page; default 10
      defaultPageSize = 10,
      defaultSortOrder = "desc",
      # reference col which table is sorted by
      defaultSorted = "sev_day_icu_perc_occup",
      # Settings in dataui examples, not sure if necessary
      defaultColDef = colDef(
            align = "center",
            headerStyle = "align-self: flex-end; font-weight:normal;"
      ),
      rowStyle = list(
            alignItems = "center",
            # add back border here
            borderBottom = "1px solid lightgray"
      ),
      # highlight rows on hover
      highlight = TRUE,
      # row expansion function
      details = row_details,
      theme = react_theme,
      columns = list(
            hospital_name = colDef(
                  name = "Hospital",
                  maxWidth = 300
            ),
            county_name = colDef(
                  name = "County",
                  maxWidth = 300
            ),
            sev_day_icu_perc_occup = colDef(
                  header = with_tooltip("ICU Occupancy", react_data_date),
                  minWidth = 100,
                  format = colFormat(percent = TRUE, digits = 1),
                  style = function(value){
                        col <- perc_pal(value)
                        if (is.na(value) | value < 0.45) {
                              list(background = col,
                                   color = "#000000")
                        } else {
                              list(background = col)
                        }
                  }
            ),
            sev_day_hosp_perc_occup = colDef(
                  header = with_tooltip("Hospital Bed Occupancy", react_data_date),
                  maxWidth = 110,
                  format = colFormat(percent = TRUE, digits = 1),
                  style = function(value){
                        col <- perc_pal(value)
                        if (is.na(value) | value < 0.45) {
                              list(background = col,
                                   color = "#000000")
                        } else {
                              list(background = col)
                        }
                  }
            ),
            avgCovIcuTenKList = icuBeds_spark(),
            avgCovHospTenKList = hospBeds_spark()
      )
) %>% 
      dui_add_reactable_dep()

hosp_react_tab




```

\
This table shows capacity measures and counts of COVID-19 patients for hospitals at the local level. HHS publishes this [dataset](https://healthdata.gov/dataset/covid-19-reported-patient-impact-and-hospital-capacity-facility) every Monday, and the latest data should cover the interval from the Friday two-weeks prior to the previous Thursday. For example, data published on December 21st will contain data from December 11th to December 17th. Given the weekly reporting schedule and the lag between the publishing and collection, I wouldn't recommend basing any personal actions solely on these data.  

Explanations for empty cells in the table and details on my methodology of handling of missing and obscured data by the CDC can be found in the [changelog](https://github.com/ercbk/Indiana-COVID-19-Tracker/blob/master/NEWS.md).  

   * *ICU Occupancy* is the proportion of the seven-day average of occupied ICU beds to the seven-day average of available ICU beds. The average daily percent of ICU beds being used over the current seven-day period. Hovering over the column name displays data date.  
   * *Hospital Bed Occupancy* is the proportion of the seven-day average of occupied hospital beds to the seven-day average of available hospital beds. The average daily percent of hospital beds being used over the current seven-day period. Hovering over the column name displays the data date.  
   * *Average Daily ICU Beds Occupied by COVID Patients* is the average daily number of ICU beds being used by confirmed and suspected COVID-19 patients over that period. Hovering over a bar displays the average count and end date for that weekly period.  
   * *Average Daily Hospital Beds Occupied by COVID Patients* is the average daily number of hospital beds being used by confirmed and suspected COVID-19 patients over that period. Hovering over a bar displays the average count and end date for that weekly period.  
   
The table is sorted by ICU Occupancy by default, but it can also be sorted alphabetically by Hospital or County and numerically by Hospital Occupancy. The search box in the upper-right corner can be used to filter the table by Hospital or County. Clicking on the triangles on the left-side of the table will expand the row and reveal an embedded Google Map centered on that hospital (may take a second to load).  

\
\
\


