---
output: 
   html_document:
      toc: true
      toc_float: true
      css: static-style.css
---


```{r, setup, include=FALSE}

knitr::opts_chunk$set(echo=FALSE, out.width="100%", out.height="100%")

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)

# get plot paths, names, and dates
png_files <- tibble::tibble(paths = fs::dir_ls(glue::glue("{rprojroot::find_rstudio_root_file()}/Indiana-COVID-19-Tracker/plots"))) %>% 
      mutate(
            chart = stringr::str_extract(paths,
                                         pattern = "[a-z]*-[a-z]*-[a-z]*"),
            date = stringr::str_extract(paths,
                                         pattern = "[0-9]{4}-[0-9]{2}-[0-9]{2}") %>%
                  as.Date())


# render the charts with latest data
png_dates <- png_files %>% 
      group_by(chart) %>% 
      summarize(newest_date = max(date))

ind_combo_date <-  png_dates$newest_date[[6]]
pos_policy_date <-  png_dates$newest_date[[8]]
goog_mob_date <-  png_dates$newest_date[[4]]
region_dea_date <-  png_dates$newest_date[[10]]
region_pos_date <-  png_dates$newest_date[[11]]
county_pos_date <-  png_dates$newest_date[[1]]
daily_re_date <-  png_dates$newest_date[[2]]
pos_rate_date <-  png_dates$newest_date[[9]]
ot_rest_date <-  png_dates$newest_date[[7]]
hosp_iv_date <-  png_dates$newest_date[[5]]
exc_death_date <-  png_dates$newest_date[[3]]

```


![Noon Update](https://github.com/ercbk/Indiana-COVID-19-Tracker/workflows/Noon%20Update/badge.svg)
![Evening Update](https://github.com/ercbk/Indiana-COVID-19-Tracker/workflows/Evening%20Update/badge.svg)

Charts updated between noon and 1pm ET and at 6:30 pm ET  
Changelog located in [NEWS.md](https://github.com/ercbk/Indiana-COVID-19-Tracker/blob/master/NEWS.md)  
\
\
\

## **Cumulative Cases and Cumulative Deaths ** 
```{r}
ind_line_path <- glue::glue("Indiana-COVID-19-Tracker/plots/ind-combo-line-{ind_combo_date}.png")
knitr::include_graphics(ind_line_path)
``` 


   * Each point is the total cases that have accumulated until that date. The lines show the exponential nature of the spread of the virus at the beginning of the epidemic and changes in trend as policy and behavior changes.  
   * *Doubling time* is the amount of time it would take for the current total to double. It assumes that the rate of change is constant. The doubling time shown here assumes that the total change from yesterday to today (i.e. single day *growth rate*) will remain constant in the future. It's not a valid assumption in this situation, but it's interesting as a hypothetical.  
\
\
\

## **Daily Positive Cases vs Cumulative Cases**  
```{r ,align="center"}
pos_policy_path <- glue::glue("Indiana-COVID-19-Tracker/plots/pos-policy-line-{pos_policy_date}.png")
knitr::include_graphics(pos_policy_path)
```
  
   * This chart shows the number of daily positive cases on the y-axis versus the cumulative total of positive cases on the x-axis.  
   * In the AEI reopening guidelines, one of the triggers for returning to stage one is five or more consecutive days where the value of daily cases has increased.  
   * Details on the various stages of re-opening can be found at [Back-on-Track Indiana](https://backontrack.in.gov/2348.htm).  
\
\
\

## **Hospitalizations, ICU Beds and Ventilator Availability**  
```{r}
hosp_iv_path <- glue::glue("Indiana-COVID-19-Tracker/plots/hosp-icu-vent-{hosp_iv_date}.png")
knitr::include_graphics(hosp_iv_path)
```

   * Even though Governor Holcomb hasn't explicitly stated the conditions that would necessitate a reversal of course to a previous stage, he did mention some benchmarks when discussing his guiding principles in his re-opening [speech](https://backontrack.in.gov/2362.htm). By negating a few of these, we can infer what some of the snapback conditions might entail.
      * An increase for 14 consecutive days in the daily total of those hospitalized for COVID-19.  
      * The proportion of available ICU beds to total ICU beds being below 40% for 14 consecutive days.
      * The proportion of available ventilators to total ventilators being below 70% for 14 consecutive days.  
\
\
\

## **Daily Positive Test Rate ** 
```{r}
pos_rate_path <- glue::glue("Indiana-COVID-19-Tracker/plots/pos-rate-line-{pos_rate_date}.png")
knitr::include_graphics(pos_rate_path)
```

   * The Positive Test Rate is the number of positive test results divided by the number of tests administered over a period of time.
   * According to the World Health Organization (WHO), the positive test rate can be a good indicator of whether a virus testing program is sufficient. It finds that successful programs have a positive test rate in the range of 3% to 12%.  
   * A Johns Hopkins [article](https://coronavirus.jhu.edu/testing/testing-positivity) calculates this rate for states over three day windows and uses the WHO benchmark to classify state programs as adequate or not.  
   * Indiana Data Hub tends to revise its counts as it continues to collect more data. When analyzing this chart, it would be prudent not to assess the last few rates too confidently as they are likely to change.  
   * The green, horizontal bar shows the WHO recommended range from 3% to 12%.  
   * U.S. average calculated using data from [The COVID Tracking Project](https://covidtracking.com/api).  
\
\
\

## **OpenTable Year-Over-Year Seated Diners **  
```{r}
ot_rest_path <- glue::glue("Indiana-COVID-19-Tracker/plots/ot-rest-line-{ot_rest_date}.png")
knitr::include_graphics(ot_rest_path)
```

   * OpenTable is a service for reviewing and making reservations at restaurants. On its "The state of the restaurant industry" [page](https://www.opentable.com/state-of-industry), it provides a couple different datasets that use reservation and food order data to summarize the state of the restaurant industry regionally by city, state, and country.  
   * The dataset being visualized uses reservation data from a sample of restaurants across Indiana. Each data point is the median daily percent difference in seated diners from the previous year to this year. So, if the day is the first Tuesday in June 2020, then the comparison is between that day and the first Tuesday of June 2019.  
   * At stages two and three, restaurants are allowed to reopen at 50% capacity. So, if a restaurant was fully-booked on a specific day in 2019 and fully booked on that same day in 2020, then we'd still see a 50% difference in seated diners for that one restaurant during stages two and three. In stage four, restaurants can open to 75% capacity, so in the same scenario above, the fully-booked, percent difference benchmark becomes 25%. Numbers around these benchmarks could indicate the level at which hoosiers are responding after each reopening stage.
   * Weekends are shown as vertical bars, and for comparison, the horizontal dashed line represents Indiana's current percent difference.  
\
\
\

## **Google Maps Mobility Indices**  
```{r}
goog_mob_path <- glue::glue("Indiana-COVID-19-Tracker/plots/goog-mob-line-{goog_mob_date}.png")
knitr::include_graphics(goog_mob_path)
```

   * Google like Apple has released some of the [data](https://www.google.com/covid19/mobility/) it collects from its Maps app that it uses to produce Community Mobility Reports and product recommendations.  
   * From the "About this Data" section in the Report:
      * These reports show how visits and length of
stay at different places change compared to
a baseline.
      * The baseline is the median value, for the
corresponding day of the week, during the 5-
week period Jan 3–Feb 6, 2020.  
   * Weekends are shown as vertical bars.  
\
\
\

## **Regional Cumulative Deaths and Cumulative Positive Cases**  
```{r, align="center"}
region_dea_path <- glue::glue("Indiana-COVID-19-Tracker/plots/region-dea-line-{region_dea_date}.png")
knitr::include_graphics(region_dea_path)
```

```{r, align="center"}
region_pos_path <- glue::glue("Indiana-COVID-19-Tracker/plots/region-pos-line-{region_pos_date}.png")
knitr::include_graphics(region_pos_path)
```

   * The y-axis has been transformed to the logarithmic base-10 scale. Slopes and trends of straight(-ish) lines are easier to compare than exponential curves.  
   * 7-day averages *greater* than 14 day averages indicate an *increasing* trend of positive cases or deaths.  
   * 7-day averages *less* than 14 day averages indicate a *decreasing* trend in positive cases or deaths.  
   * State policy [data](https://docs.google.com/spreadsheets/d/1zu9qEWI8PsOI_i8nI_S29HDGHlIp2lfVMsGxpQ5tvAQ/edit#gid=0) comes from Julia Raifman, Kristen Nocka, et al at Boston University.  
\
\
\

## **Estimated Change in County Postive Cases ** 
```{r, align="center"}
county_pos_path <- glue::glue("Indiana-COVID-19-Tracker/plots/county-pos-bar-{county_pos_date}.png")
knitr::include_graphics(county_pos_path)
```

   * The top 20 counties according to estimated average percent change are shown. Counties at the top are experiencing the highest average growth rates in positive test results.
   * The average percent changes are estimated using county data in a log-linear model in order to capture exponential growth if an outbreak occurs.  
\
\
\

## **Effective Reproductive Number**    
```{r}
daily_re_path <- glue::glue("Indiana-COVID-19-Tracker/plots/daily-re-line-{daily_re_date}.png")
knitr::include_graphics(daily_re_path)
``` 

   * The instantaneous effective reproduction number, R~e~ is an estimate of the average number of people that an infected person is spreading the virus to for that time window.  
   * Estimates *slighly lower* than 1 indicate that the spread of the virus is decreasing. Estimates *slightly larger* than 1 indicate a slow spread of the virus. Estimates of *2 or greater* mean the virus is spreading rapidly.  
   * The shaded area shows the 80% credible interval where the true value is likely to be.
   * Further details on this metric can be found at a research [site](https://cbdrh.github.io/ozcoviz/#national-level-r_t) set-up by South Western Sydney Clinical School and the Centre for Big Data Research in Health.  
\
\
\

## **Excess Deaths**  
```{r}
exc_death_path <- glue::glue("Indiana-COVID-19-Tracker/plots/excess-death-col-{exc_death_date}.png")
knitr::include_graphics(exc_death_path)
``` 

  * The [CDC's](https://www.cdc.gov/nchs/nvss/vsrr/covid19/excess_deaths.htm) model uses data from 2015 to present to calculate estimates for weekly death counts. "Excess deaths" is defined as the difference between these estimates and the actual reported number of deaths.  
  * A portion of these excess deaths could be misclassified COVID-19 deaths, "or potentially could be indirectly related to COVID-19 (e.g., deaths from other causes occurring in the context of health care shortages or overburdened health care systems)."  
  * From the CDC site regarding the Causes-of-deaths data, "These causes were selected based on analyses of comorbid conditions reported on death certificates where COVID-19 was listed as a cause of death. Deaths with an underlying cause of death of COVID-19 are not included in these estimates of deaths due to other causes, but deaths where COVID-19 appeared on the death certificate as a multiple cause of death may be included in the cause-specific estimates."  